\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-25 paper_resources/doc/paper.tex
Polishing in the Exp. section

\end{itemize}
The above-mentioned benchmark sets may be biased as models which describe physical laws that usually impose   symmetries, periodicity w.r.t. some variables, internal separability on some variables etc. In that way, the search process and the search space can be, in-advance, significantly reduced. Thus, some methods may get significant benefit with this regard, while some others may be in the subordinate situation. 
In order to permit any kind of favoritism in the input data, we generate a set of randomly generated (non-biased) SR problem instances; this benchmark set is labelled by \textsc{Random}. These instances are generated according the following procedure. \fxnote{Opisati proceduru generisanja instanci + kombinaciju parametara za instance, itd.}

\subsection{Parameter tuning}
The \textsc{RILS}-\textsc{ROLS} algorithm employs just two parameters that need to be tuned, $penalty_{size}$ and $tolerance_{error}$. The first parameter is fixed to 0.001(?), according to the preliminary experimental evaluation we conducted. The second parameter is tuned by monitoring the two errors the algorithm produced on a randomly chosen training set of xx problem instances; among the following values \{ x, xx, ... \} we chose $tolerance_{error}=?$ for which the largest average $R^2$ and the smallest average RMSE error over all problem instances from the training set are produced. %Note that the training set is chosen randomly with a cardinalty of xx instances.

@@

\subsection{Comparison with other methods on the   ground--truth benchmark sets}
In this section, we evaluate the performance of \textsc{RILS-ROLS} algorithm and compare it to the 14 other competitors from the literature. The numerical results are reported in terms of the three Tables~\ref{tab:comp_noise0}--\ref{tab:comp_noise001}. The results for the two ground-truth benchmark sets, \textsc{Feynman} and \textsc{Storgatz}, are given w.r.t.\   the following values of the noise levels of 0.0 (no noise), 0.001 (a low-level noise), and 0.01 (a high-level noise), respectively. In details, the white Gaussian noise was added to the target ($y$-axis) values as a fraction of the signal RMS value, also applied in~\cite{la2021contemporary}, i.e. for target noise level $\alpha$, we have
$$ y_{noise} = y + \epsilon, \epsilon \sim \mathcal{N}\left(0, \alpha \sqrt{\frac{1}{n} \sum {y_i^2}}\right),$$
where $n$ relates to the number of items in the input data.

Each table consists of three blocks. The first block reports the algorithm whose performances are reported (by a row). The second block reports, with two columns,  percentages of the instances where the model which fits to the exact one has been found by respective algorithms on \textsc{Feynman} and \textsc{Strogatz} benchmark set, respectively. The third column of this block provides the overall summary exact percentage of success over all instances from the both benchmark sets. The third block displays the results of the respective algorithm in terms of the $R^2$ score which indicate on  accuracy, i.e. in our cases when the algorithm  at least reached the accuracy ($R^2$ score) of 0.999.  Note that the termination of our algorithm may not always be triggered by reaching an appropriate $R^2$ score of 0.999, but also due to the reached time limit. These results are of interest as when compared to corresponding exact percentage it may point out the existence of an over-fitting. This block consists of three rows that indicate the percentage of instances where respective algorithm is able to reach the desired accuracy (fulfilling the condition $R^2 \geq 0.999$ upon termination) on the both real--world benchmark sets separately, and the overall percentage of success rate (over all considered instances) w.r.t. the same criterion of reaching the desired accuracy.

@@

In short, the exact results, thus the number of instances (in percentages) for which the output of respective algorithm match to the exact model, are displayed in the second block of each of the tables. The numbers of instances (in percentages) for which each algorithms deliver a precise enough model (i.e. satisfying $R^2\geq 0.999$ upon its termination) are displayed in the third block of each of the tables.

>>>>

\end{itemize}
The above-mentioned benchmark sets may be biased since the models which describe physical laws usually impose   symmetries, periodicity w.r.t. some variables, internal separability on some variables etc. In that way, the search process and the search space can be pre-processed and thus significantly reduced. Thereby, some methods may get significant benefit with this regard, while some others may be in a subordinate situation. 
In order to permit any kind of favoritism in the input data, we generate a set of randomly generated (non-biased) SR problem instances; this benchmark set is labelled by \textsc{Random}. These instances are generated according the following procedure. \fxnote{Opisati proceduru generisanja instanci + kombinaciju parametara za instance, itd.}

\subsection{Parameter tuning}
The \textsc{RILS}-\textsc{ROLS} algorithm employs two parameters that need to be tuned, $penalty_{size}$ and $tolerance_{error}$. The first parameter is fixed to 0.001(?), according to the preliminary experimental evaluation. The second parameter is tuned by monitoring the two errors the algorithm produced on a randomly chosen training set of xx problem instances. Among the following values \{ x, xx, ... \} we chose $tolerance_{error}=?$ for which the largest average $R^2$ and the smallest average RMSE error on all problem instances from the training set are produced. %Note that the training set is chosen randomly with a cardinalty of xx instances.

@@

\subsection{Comparison with other methods on the   ground--truth benchmark sets}
In this section, we evaluate the performance of \textsc{RILS-ROLS} algorithm and compare it to the 13 other competitors from the literature. The numerical results are reported in terms of the three Tables~\ref{tab:comp_noise0}--\ref{tab:comp_noise001}. The results for the two ground-truth benchmark sets, \textsc{Feynman} and \textsc{Storgatz}, are given w.r.t.\   the following level of noise: 0.0 (no-noisy data), 0.001 (low-level noise), and 0.01 (high-level noise), respectively. In details, the white Gaussian noise was added to the target ($y$-axis) values as a fraction of the signal RMS value, applied also in~\cite{la2021contemporary}, i.e. for target noise level $\alpha$ we have
$$ y_{noise} = y + \epsilon, \epsilon \sim \mathcal{N}\left(0, \alpha \sqrt{\frac{1}{n} \sum {y_i^2}}\right),$$
where $n$ relates to the number of items in the input data.

Each table consists of three blocks. The first block reports the algorithm (by a row) whose performances are reported. The second block reports in two columns  percentages of the instances where the exact model has been found by respective algorithms on \textsc{Feynman} and \textsc{Strogatz} benchmark set, respectively. The third column of this block provides the overall summary exact percentage of success of the algorithm over all instances from the both benchmark sets. The third block displays the results of the respective algorithm in terms of the final $R^2$ score which indicates on accuracy of found solution -- by our criterion when the algorithm  at least reached the accuracy ($R^2$ score) of 0.999.  Note that the termination of our algorithm may not always be triggered by reaching an appropriate $R^2$ score of 0.999, but also due to the reached time limit. These results are of a particular interest as they may point out on the presence of strong over-fitting in contrast to finding the right model. This block consists of three rows that indicate the percentage of instances where respective algorithm is able to reach the desired accuracy (fulfilling the condition $R^2 \geq 0.999$ upon termination) on the both ground--truth benchmark sets separately, and the overall percentage of success rate (over all considered instances) w.r.t. the criterion of reaching the final accuracy of 0.999.

@@

In short, the exact results, thus the number of instances (in percentages) for which the output of respective algorithm is able to find the exact model, are displayed in the second block of each of the tables. The numbers of instances (in percentages) for which each algorithms deliver a precise enough model (i.e. satisfying $R^2\geq 0.999$ upon its termination) are displayed in the third block of each of the tables.
